---
title: "Spatial_Assignment_3"
author: "Group22: Jeb Polstein, Matt Khinda, Yoji Toriumi"
date: "9/16/2021"
output: html_document
---

```{r, echo=FALSE}
library(sf)
library(tigris)
library(tidyverse)
library(zonebuilder)
library(ggthemes)
library(ggspatial)
library(units)
library(nngeo)
```

## Extracted some data from [Analyze Boston](https://data.boston.gov/)
```{r Import Data, echo=FALSE}
hospitals <- read_sf("https://bostonopendata-boston.opendata.arcgis.com/datasets/8f1e5741a4ee4c0f82925d17aab3c002_2.kml?outSR=%7B%22latestWkid%22%3A2249%2C%22wkid%22%3A102686%7D")
universities <- read_sf("https://bostonopendata-boston.opendata.arcgis.com/datasets/cbf14bb032ef4bd38e20429f71acb61a_2.kml?outSR=%7B%22latestWkid%22%3A2249%2C%22wkid%22%3A102686%7D")
policest <- read_sf("https://bostonopendata-boston.opendata.arcgis.com/datasets/e5a0066d38ac4e2abbc7918197a4f6af_6.kml?outSR=%7B%22latestWkid%22%3A2249%2C%22wkid%22%3A102686%7D")
openspace <- read_sf("https://bostonopendata-boston.opendata.arcgis.com/datasets/2868d370c55d4d458d4ae2224ef8cddd_7.kml")
bikenet <- read_sf("https://bostonopendata-boston.opendata.arcgis.com/datasets/1f63e8e31f664833827c407c021ad60c_0.kml?outSR=%7B%22latestWkid%22%3A2249%2C%22wkid%22%3A102686%7D")
neighbor <- read_sf("https://bostonopendata-boston.opendata.arcgis.com/datasets/3525b0ee6e6b427f9aab5d0a1d0a1a28_0.kml?outSR=%7B%22latestWkid%22%3A2249%2C%22wkid%22%3A102686%7D")
hydro <- read_sf("https://bostonopendata-boston.opendata.arcgis.com/datasets/2b3c0fa13b1c468eb702a3645fcc6bcb_5.kml")
tracts <- tracts(state = "MA", county = "Suffolk") 
boundary <- st_read("https://bostonopendata-boston.opendata.arcgis.com/datasets/142500a77e2a4dbeb94a86f7e0b568bc_9.geojson?outSR=%7B%22latestWkid%22%3A2249%2C%22wkid%22%3A102686%7D", quiet = TRUE)
```


```{r Transforming Data, echo=FALSE}
MA_state_plane <- "+proj=lcc +lat_1=41.71666666666667 +lat_2=42.68333333333333 +lat_0=41 +lon_0=-71.5 +x_0=200000 +y_0=750000 +ellps=GRS80 +units=m +no_defs"
hospitals <- hospitals %>% 
  st_transform(2249)
universities <- universities %>% 
  st_transform(2249)
policest <- policest %>% 
  st_transform(2249)
openspace <- openspace %>% 
  st_transform(2249)
bikenet <- bikenet %>% 
  st_transform(2249)
neighbor <- neighbor %>% 
  st_transform(2249)
hydro <- hydro %>% 
  st_transform(2249)
tracts <- tracts %>% 
  st_transform(2249)
boundary <- boundary %>% 
  st_transform(2249)
tracts <- tracts[boundary, ]
boundary <- boundary %>% st_transform(2249)
```

## FYI, all of the data are shown as follows
```{r, echo=FALSE}
ggplot(neighbor) +
  geom_sf() +
  geom_sf(data = hospitals, color = "red") +
  geom_sf(data = universities, color = "orange") +
  geom_sf(data = policest, color = "green") +
  geom_sf(data = bikenet) +
  geom_sf(data = openspace) +
  geom_sf(data = hydro, fill = "blue") +
  theme_classic() +
  annotation_scale()
```


##
```{r, echo=FALSE}
openspace_buffer <- st_buffer(openspace, dist = 100) %>% 
  st_union()

ggplot(openspace_buffer) +
  geom_sf() +
  theme_map()
```

```{r, echo=FALSE}
hospitals_openspace <- hospitals[openspace_buffer, ]

ggplot(openspace_buffer) +
  geom_sf() +
    geom_sf(data = hospitals_openspace,
          color = "tomato",
          size = 0.5)
  theme_map()
```

```{r, echo=FALSE}
hospitals <- hospitals %>% 
  st_join(hospitals_openspace) %>% 
  mutate(by_openspace = !is.na(Name.y))
```

## how many hospitals are within 100m of openspaces
```{r, echo=FALSE}
n_openspace_hospitals <- sum(hospitals$by_openspace)

n_openspace_hospitals
```
## ratio
```{r, echo=FALSE}
n_hospitals <- length(hospitals$by_openspace)

pct_openspace_hospitals <- n_openspace_hospitals / n_hospitals * 100

pct_openspace_hospitals
```

```{r, echo=FALSE}
left_side <- st_bbox(hospitals)$xmin
top_side  <- st_bbox(hospitals)$ymax

ggplot(hydro) +
  geom_sf(fill = "cadetblue1", color = NA) +
  geom_sf(data = hospitals, size = 3,
          aes(color = by_openspace)) +
  scale_color_manual(values = c("chartreuse1", "tomato"),
                     name = "Boston Hospitals\nby distance to openspace",
                     labels = c("No openspace within 100m",
                                "Openspace within 100m")) +
  annotation_scale(location = "br") +
  annotation_north_arrow(location = "tr",
                         style = north_arrow_minimal()) +
  annotate(geom =  "text", x = left_side,
           y = top_side,
           label = paste("Of the ", 
                         prettyNum(n_hospitals),
                         " hospitals in Boston\n", 
                         prettyNum(n_openspace_hospitals),
                         " (", 
                         prettyNum(pct_openspace_hospitals, digits = 0),
                         "%) are within 100\nmeters of openspace.",
                         sep = ""),
           hjust = 0, vjust = 0, size = 3) +
  theme_map() +
    theme(panel.background = element_rect(fill = "antiquewhite"),
        legend.background = element_rect(fill = alpha("white", 0.5), 
                                         color = "gray"))
```

```{r, echo=FALSE}
neighbor <- neighbor %>% 
  mutate(num_police = lengths(st_covers(neighbor, policest)))

ggplot(neighbor) +
  geom_sf(color = NA,
          aes(fill = num_police)) +
  scale_fill_viridis_c(name = "Boston neighborhoods\nbynumber of police station",
                       breaks = breaks <- seq(0, 2, by = 1),
                       labels = paste(prettyNum(breaks),
                                      "police stations")) +
  annotation_scale(location = "br") +
  annotation_north_arrow(location = "tr",
                         style = north_arrow_minimal()) +
  theme_map() +
  theme(legend.background = element_rect(fill = alpha("white", 0.5), 
                                         color = "gray"))


```

```{r, echo=FALSE}
neighbor <- neighbor %>%
  mutate(area = set_units(st_area(neighbor), km^2)) %>%
  mutate(police_dens = as.numeric(num_police/ area))

ggplot(neighbor) +
  geom_sf(color = NA,
          aes(fill = police_dens)) +
  scale_fill_viridis_c(name =
                         "Boston neighborhoods \nby police station density",
                       breaks = breaks <- seq(0, 0.62, by = 0.1),
                       labels = paste(prettyNum(breaks),
                                      "police stations per sq km")) +
  annotation_scale(location = "br") +
  annotation_north_arrow(location = "tr",
                         style = north_arrow_minimal()) +
  theme_map() +
  theme(legend.position = "right",
         legend.background = element_rect(fill = alpha("white", 0.5), 
                                         color = "gray"))
```

##Closest Point

```{r, echo=FALSE}
openspace <- openspace %>% 
  mutate(hospitals_dist = st_nn(openspace, hospitals,
                                returnDist = TRUE)$dist) %>% 
  mutate(hospitals_dist = as.numeric(hospitals_dist))
```

```{r, echo=FALSE}
avg_hospitals_dist <- mean(openspace$hospitals_dist)

avg_hospitals_dist
```

```{r, echo=FALSE}
right_side <- st_bbox(openspace)$xmax
left_side  <- st_bbox(openspace)$xmin
top_side <- st_bbox(openspace)$ymax
bottom_side <- st_bbox(openspace)$ymin

ggplot(hydro) +
  geom_sf(fill = "cadetblue1", color = NA) +
  geom_sf(data = openspace, aes(fill = hospitals_dist)) +
  coord_sf(xlim = c(left_side, right_side),
           ylim = c(bottom_side, top_side), expand = FALSE) +
  scale_fill_viridis_c(name = "Boston Openspace \nby distance to a hospital") +
  annotation_scale(location = "br") +
  annotation_north_arrow(location = "tr",
                         style = north_arrow_nautical()) +
  annotate(geom = "text", x = left_side + 300,
           y = top_side - 500,
           label = paste("On average, a Boston openspace\nis ",
                         prettyNum(avg_hospitals_dist, digits = 3),
                         "m",
                         sep = ""),
           hjust = 0, size = 3) +
  theme_map() +
  theme(panel.background = element_rect(fill = "antiquewhite"),
        legend.background = element_rect(fill = alpha("white", 0.5), 
                                         color = "gray"))
```

##Identifying Overlapping Polygons
```{r, echo=FALSE}
neighbor <- neighbor %>%
  mutate(num_hospitals = lengths(st_covers(neighbor, hospitals))) %>% 
  mutate(has_hospitals = num_hospitals > 0)

n_hospitals_neighbor <- sum(neighbor$has_hospitals)

n_hospitals_neighbor
```

```{r, echo=FALSE}
left_side  <- st_bbox(neighbor)$xmin
top_side <- st_bbox(neighbor)$ymax

ggplot(hydro) +
  geom_sf(fill = "cadetblue1", color = NA) +
  geom_sf(data = neighbor,
          aes(fill = has_hospitals)) +
  scale_fill_manual(values = c("chartreuse1", "tomato"),
                    name = "Boston Neighborhoods\nby presence of a hospital",
                    labels = c("Neighborhood without\nany presence of a hospital",
                               "Neighborhood with \n presence of a hospital(s)")) +
  annotation_scale(location = "br") +
  annotation_north_arrow(location = "tr",
                         style = north_arrow_minimal()) +
  annotate(geom = "text", x = left_side, 
           y = top_side - 1000, 
           label = paste(n_hospitals_neighbor,
                         "of Boston's",
                         length(neighbor$Name),
                         "neighborhoods contain a hospital(s)."),
           hjust = 0, vjust = -5, size = 3) +
  theme_map() +
  theme(panel.background = element_rect(fill = "antiquewhite"),
        legend.background = element_rect(fill = alpha("white", 0.5), 
                                         color = "gray"))

```

```{r, echo=FALSE}
grid <- st_sf(st_make_grid(boundary, n = c(20, 20)))
grid <- grid[boundary, ]

grid <- grid %>%
  mutate(num_hospitals = lengths(st_covers(grid, hospitals))) %>% 
  mutate(area = set_units(st_area(grid), km^2)) %>% 
  mutate(hospitals_dens = as.numeric(num_hospitals / area))


plot_5 <- ggplot(grid) +
  annotation_map_tile(zoomin = 0, progress = "none", type = "cartolight") +
  geom_sf(data = grid, alpha = 1/2, aes(fill = hospitals_dens)) +
  scale_fill_viridis_c(name = "Schools per square km") +
  labs(caption = "Map tiles and data by OpenStreetMap") +
  theme_void()

plot_5
```

```{r, echo=FALSE}
clock <- zb_zone("Boston", distance = 0.5, distance_growth = 0, n_circles = 20)
clock <- clock %>% st_transform(2249)
clock <- clock[boundary, ]
clock <- clock %>%
  mutate(num_hospitals = lengths(st_covers(clock, hospitals))) %>% 
  mutate(area = set_units(st_area(clock), km^2)) %>% 
  mutate(hospitals_dens = as.numeric(num_hospitals / area))

plot_7 <- ggplot(clock) +
  annotation_map_tile(zoomin = 0, progress = "none", type = "cartolight") +
  geom_sf(alpha = 1/2, aes(fill = hospitals_dens)) +
  scale_fill_viridis_c(name = "Hospitals per square km") +
  labs(caption = "Map tiles and data by OpenStreetMap") +
  theme_void()

plot_7
```

##Matt's Explorations  

####Open spaces that are within a 400 meters of a body of water 
This map looks at which open spaces in Boston are within 400 meters of a body of water. The 400 meter buffer was chosen because it is often used as the normative planning standard for a 5 minute walk. This calculation was done by trimming the open spaces dataset to remove parks that extended beyond the limits of the water dataset, adding a 400m buffer to the water polygons, then calculating the intersection of the open space polygons and buffered water polygons. For this assignment, this map serves to satisfy calculation 14 (the number and proportion of C polygons that overlap with D polygons).

```{r matt plot 1, echo=FALSE}
#trimming open space dataset to remove Brewster Islands that fall outside the water polygon's xlim
bstn_openspace <- openspace[boundary, ]

#creating a 400m buffer from the water
hydro_buffer <- st_buffer(hydro, dist = 400) %>%
  st_union()

#calculating overlap in open space and water + buffer
openspace_water <- bstn_openspace %>%
  mutate(num_water = lengths(st_intersects(bstn_openspace, hydro_buffer))) %>%
  mutate(has_water = num_water > 0)

n_openspace_water <- sum(openspace_water$has_water)
n_openspace <- nrow(bstn_openspace)
pct_openspace_water <- n_openspace_water / n_openspace * 100

#plotting
left_side  <- st_bbox(hydro)$xmin
bottom_side <- st_bbox(hydro)$ymin

ggplot() +
  geom_sf(data = hydro, color = NA, fill = "lightblue") +
  geom_sf(data = openspace_water, color = NA, aes(fill = has_water), alpha = 0.4, size = 0.5) +
  scale_fill_manual(values = c("red", "darkgreen"), 
                    name = "Boston open spaces\nby distance from a body of water", 
                    labels = c("Open spaces that are not within\n 400m of a body of water", 
                               "Open spaces that are within\n 400m of a body of water"),
                    ) +
  
  geom_sf(data = hydro_buffer, aes(color = "400m from waterfront", linetype = "400m from waterfront"), alpha = 0.8, fill = NA) +
  scale_color_manual(values = "darkgrey", name = "") +
  scale_linetype_manual(values = "longdash", name = "") +
  
  annotate(geom = "text", x = left_side, 
           y = bottom_side - 5000, 
           label = paste(n_openspace_water,"(",
                          round(pct_openspace_water, digits = 1) ,
                         "%) of Boston's open spaces are \nwithin 400 meters of", 
                         "a body of water."),
           hjust = 0, vjust = 0, size = 3) +
  
  annotation_scale(location = "br") +
  annotation_north_arrow(location = "tr", style = north_arrow_minimal()) +
  
  labs(caption = "Data from data.boston.gov") +
  theme_map() +
  theme(legend.position = c(0, 0.7),
        legend.background = element_rect(fill = alpha("white", 0.5)))

```
  
####Hospitals with a 1000 meter radius of Boston University buildings
This map looks at which hospitals in Boston are within 1000 meters of a Boston University building. This calculation was done by filtering the universities only to include Boston University, adding a 1000m buffer to the BU building points, then subsetting the hospital dataset only to include those within the buffer. For this assignment, this map serves to satisfy calculation 2 (The number and proportion of B points with a specified distance of A points).

```{r matt plot 2, echo=FALSE}
#filtering for BU buildings
BU_universities <- universities %>%
  filter(grepl("Boston University", Name))

#adding a buffer to BU buildings
BU_buffer <- st_buffer(BU_universities, dist = 1000) %>%
  st_union

#subsetting hospitas within the buffer
BU_hospitals <- hospitals[BU_buffer, ]

#calculating number and percent of hospitals near BU
n_BU_hospitals <- nrow(BU_hospitals)
pct_BU_hospitals <- n_BU_hospitals / n_hospitals * 100

#plotting
left_side  <- st_bbox(hydro)$xmin
top_side <- st_bbox(hydro)$ymax

ggplot() +
  geom_sf(data = hydro, color = NA, fill = "lightblue") +
  geom_sf(data = boundary, fill = "white", color = NA) +
  geom_sf(data = BU_buffer, fill = NA, color = "black", size = .5, linetype = "twodash") +
  geom_sf(data = hospitals, aes(color = "Hospitals more than 1000 meters from BU building"), size = .5) +
  geom_sf(data = BU_hospitals, aes(color = "Hospitals within 1000 meters of a BU building"), size = .5) +
  geom_sf(data = BU_universities, aes(color = "BU building"), size = .5) +
  scale_color_manual(values = c("black","purple","red"),
          name = "Hospitals by distance to \nBoston University Buildings") +
  
  annotate(geom = "text", x = left_side, 
           y = top_side - 4000, 
           label = paste("Of the ", 
                         prettyNum(n_hospitals, big.mark = ","),
                         " hospitals in Boston\n", 
                         prettyNum(n_BU_hospitals, big.mark = ","),
                         " (", 
                         prettyNum(pct_BU_hospitals, digits = 0),
                         "%) are within 1000\nmeters of a Boston University building.",
                         sep = ""),
           hjust = 0, vjust = 0, size = 3) +
  
  annotation_scale(location = "br") +
  annotation_north_arrow(location = "tr", style = north_arrow_minimal()) +
  
  labs(caption = "Data from data.boston.gov") +
  
  theme_map() +
  theme(panel.background = element_rect(fill = "#EDEDED"),
        legend.background = element_rect(fill = alpha("white", 0.5), 
                                         color = "black", size = 0.25))

```

```{r, echo=FALSE}

```

```{r, echo=FALSE}

```

```{r, echo=FALSE}

```

```{r, echo=FALSE}

```










